---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![R-CMD-check](https://github.com/IDEMSInternational/outfillingR/workflows/R-CMD-check/badge.svg)](https://github.com/IDEMSInternational/outfillingR/actions)
[![Codecov test coverage](https://codecov.io/gh/IDEMSInternational/outfillingR/branch/main/graph/badge.svg)](https://app.codecov.io/gh/IDEMSInternational/outfillingR?branch=main)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-green.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![Project Status: WIP – Initial development is in progress, but there
has not yet been a stable, usable release suitable for the
public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![CRAN
status](https://www.r-pkg.org/badges/version/outfillingR)](https://CRAN.R-project.org/package=outfillingR)
[![license](https://img.shields.io/badge/license-LGPL%20(%3E=%203)-lightgrey.svg)](https://www.gnu.org/licenses/lgpl-3.0.en.html)
<!-- badges: end -->

## Overview

`outfillingR` is an R package designed to address missing data through infilling and outfilling methods. The package is particularly suited for environmental and meteorological data, including rainfall datasets, and allows users to incorporate statistical distributions, custom bins, and Markov chain techniques into their workflows.

## Installation

Install the development version from GitHub with:

```{r, eval = FALSE}
#install.packages("devtools")
devtools::install_github("IDEMSInternational/outfillingR")
```

## Examples

Here is an example of infilling data.

```{r, message = FALSE, warning = FALSE}
# Load the outfillingR library
library(outfillingR)
```

```{r, message = FALSE, warning = FALSE, eval=FALSE, include=TRUE}
# Load example dataset included with the package
data("zambia_data") 

# Explore the dataset
head(zambia_data)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
# Load example dataset included with the package
data("zambia_data") 

# Explore the dataset
head(zambia_data) %>%
  knitr::kable()
```

```{r, message = FALSE, warning = FALSE, eval=FALSE, include=TRUE}
# Perform infilling using default parameters
infilled_data <- do_infilling(
  data = zambia_data,
  station = "Station_name",
  date = "date",
  rainfall = "rainfall",
  rfe = "rfe",
  lon = "lon",
  lat = "lat",
  station_to_exclude = "Station A",
  rainfall_estimate_column = "rfe",
  custom_bins = c(1, 5, 10, 15, 20),
  count_filter = 5,
  min_rainy_days_threshold = 30,
  target_months = c(5, 6, 7, 8, 9),
  distribution_flag = "gamma",
  markovflag = TRUE
)

# View the results
head(infilled_data)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
# Perform infilling using default parameters
infilled_data <- do_infilling(
  data = zambia_data,
  station = "Station_name",
  date = "date",
  rainfall = "rainfall",
  rfe = "rfe",
  lon = "lon",
  lat = "lat",
  station_to_exclude = "Station A",
  rainfall_estimate_column = "rfe",
  custom_bins = c(1, 5, 10, 15, 20),
  count_filter = 5,
  min_rainy_days_threshold = 30,
  target_months = c(5, 6, 7, 8, 9),
  distribution_flag = "gamma",
  markovflag = TRUE
)

# View the results
head(infilled_data)
```

## Contributing
We welcome contributions! If you'd like to contribute, please submit a pull request or file an issue on the GitHub repository.